---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: https://github.com/peachvegetable/STA302-mini-essay-10"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(knitr)
library(rstanarm)
library(modelsummary)
library(boot)
library(broom.mixed)
```


# Introduction

Generates a model on the dataset obtained from kaggle Student Study Performance, how math score is related to various variables.



# Data {#sec-data}

```{r}
#| label: tbl-meanvariance
#| tbl-cap: mean and variance of the math score
#| echo: false
#| warning: false

analysis_data <- read.csv("../data/analysis_data/analysis_data.csv")


mean_var <- analysis_data|>
  summarise(
    Mean_Deaths = mean(math_score, na.rm = TRUE),
    Variance_Deaths = var(math_score, na.rm = TRUE)
  )

# Use kable to create a table
kable_table <- kable(mean_var, format = "latex", digits = c(1, 1),
                     col.names = c("Mean", "Variance"))

kable_table

```

```{r}
#| label: tbl-meanvariance
#| fig-cap: mean and variance of the math score
#| echo: false
#| warning: false

neg_binomial_model <- readRDS("../models/neg_binomial_model.rds")

modelsummary(
  list(
    "Negative binomial" = neg_binomial_model
    )
)
```




# Model


## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.  

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}


\newpage


# References


